---
title: "Fig 1 - OLS Simulation"
author: Casey Breen 
---

Summary: This .Rmd file contains code to replicate Figure 1 in "Mortality Modeling of Partially Observed Cohorts Using Administrative Death Records." 

Description of Figure 1: The ratio of truncated to untruncated regression statistics: standard errors (panel a), regression coefficients (panel b), and t-statistics (panel c) for different truncation windows. A value of 1 corresponds to the truncated and untruncated statistic having the same value; values less than 1 correspond to the truncated statistic being smaller than the untruncated statistic.

```{r}
## library packages
library(tidyverse)
library(gompertztrunc)
library(cowplot)
library(broom)
library(here)

## Set
set.seed(5)
```


```{r}
## create list to store results
results <- list()
i <- 1

## Simulate difference truncation windows, systematically varying the start year and the width of the window
for (start_year in c(1865:1880)) {
  for (window_width in c(1:30)) {
    ## smulate death distribution
    sim_data <- gompertztrunc::gompertztrunc_simu(n = 200000, formula = death_age ~ sex, coefs = c("sex" = -0.2), dummy = c(TRUE)) %>%
      mutate(byear = 1800) %>%
      mutate(dyear = byear + floor(death_age))

    ## filter death distribution to be within window
    sim_data_trunc <- sim_data %>%
      filter(dyear >= start_year & dyear <= start_year + window_width)

    ## estimate adjustment factor
    adj_factor <- 200000 / nrow(sim_data_trunc)

    ## simulate gompertz distribution
    sim_data <- gompertztrunc::gompertztrunc_simu(n = adj_factor * 200000, formula = death_age ~ sex, coefs = c("sex" = -0.2), dummy = c(TRUE)) %>%
      mutate(byear = 1800) %>%
      mutate(dyear = byear + floor(death_age))

    ## truncate data
    sim_data_trunc <- sim_data %>%
      filter(dyear >= start_year & dyear <= start_year + window_width)

    ## fit linear model
    lm_model <- lm(death_age ~ sex, data = sim_data_trunc)

    ## store results
    results[[i]] <- tibble(
      coef = summary(lm_model)$coefficients["sex", "Estimate"],
      se = summary(lm_model)$coefficients["sex", "Std. Error"],
      start_year = start_year,
      window_width = window_width,
      n = nrow(sim_data_trunc)
    )


    i <- i + 1
    cat(i)
  }
}

## simulate gompertz window
sim_data <- gompertztrunc::gompertztrunc_simu(n = 365000, formula = death_age ~ sex, coefs = c("sex" = -0.2), dummy = c(TRUE)) %>%
  mutate(byear = 1800) %>%
  mutate(dyear = byear + floor(death_age))

## filter to death ages over 65
sim_data <- sim_data %>%
  filter(death_age > 65)

## rerun second model
lm_model2 <- lm(death_age ~ sex, data = sim_data)

## estimate coefficients of the linear model
untruncated_regression <- tibble(
  coef = summary(lm_model2)$coefficients["sex", "Estimate"],
  se = summary(lm_model2)$coefficients["sex", "Std. Error"],
  start_year = start_year,
  window_width = window_width,
  n = nrow(sim_data_trunc),
  t_stat = coef / se
)

##
data_forplot <- bind_rows(results) %>%
  ungroup() %>%
  mutate(t = coef/se) %>% 
  mutate(
    se = se / untruncated_regression$se,
    t = t / untruncated_regression$t_stat,
    coef = coef / 0.986
  )

## plot standard errors
plota <- ggplot(data_forplot, aes(x = start_year - 1800, y = window_width)) +
  geom_tile(aes(fill = se)) +
  scale_fill_viridis_c(option = "magma") +
  cowplot::theme_minimal_hgrid(15) +
  labs(
    y = "Width of Window (Years)",
    x = "Starting Age",
    fill = "SE / SE (untruncated)",
    title = "Standard Errror",
    caption = ""
  ) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm")
  )

## plot coefficients
plotb <- ggplot(data_forplot, aes(x = start_year - 1800, y = window_width)) +
  geom_tile(aes(fill = coef)) +
  scale_fill_viridis_c(option = "magma") +
  cowplot::theme_minimal_hgrid(15) +
  labs(
    y = "Width of Window (Years)",
    x = "Starting Age",
    fill = "Coef / Coef (untruncated)",
    title = "Regression Coefficient"
  ) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm")
  )

## plot t-statistic
plotc <- ggplot(data_forplot, aes(x = start_year - 1800, y = window_width)) +
  geom_tile(aes(fill = t)) +
  scale_fill_viridis_c(option = "magma") +
  cowplot::theme_minimal_hgrid(15) +
  labs(
    y = "Width of Window (Years)",
    x = "Starting Age",
    fill = "T-Stat / T-Stat (untruncated)",
    title = "T-Statistic"
  ) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm")
  )

## create combined plot
plot_sim <- cowplot::plot_grid(plotb, plota, plotc, ncol = 1, labels = "auto")

title <- ggdraw() +
  draw_label(
    "OLS under Double Truncation",
    fontface = "bold",
    x = 0,
    hjust = 0,
    vjust = 2.5,
    size = 22
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 0)
  )

plot_sim_title <- plot_grid(
  title, plot_sim,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1)
)

## save plot
ggsave(plot_sim_title, filename = here("figures/fig1-simulated_window_widths.png"), height = 17, width = 10)
```
