---
title: "Figure 2 - Regression Example"
author: Joshua R. Goldstein 
---

Summary: This .Rmd file contains code to replicate Figure 2 in "Mortality Modeling of Partially Observed Cohorts Using Administrative Death Records." 

Description of Figure 2: Simulated example of regression on effect of covariate on death age based on full data (left) and truncated death ages only (right). The truncation reduces the estimated ``effect'' by about 2/3, from 0.45 to 0.15, in this example. Note: Simulation is sample of 100 individuals following  $y = 30 + 0.5x + \epsilon$, with $x \sim N(\mu = 100, \sigma = 5)$ and $\epsilon \sim N(\mu = 0, \sigma = 5)$.


```{r}
## library package
library(tidyverse)
library(flexsurv)
library(here)

## set seed
set.seed(9999)
```

```{r}
## set parameters
N <- 100
z <- rnorm(N, 0, 1)
b <- -.5
alpha <- 0.5 * 10^-4
beta <- 1 / 10
y <- rgompertz(N, shape = beta, rate = alpha * exp(z * b))
par(mfrow = c(1, 1))
plot(z, y, col = "grey")

## true regression line
m <- lm(y ~ z)
abline(m, col = "black")

## truncated
L <- 60
R <- 75
abline(h = c(L, R), col = "grey")
mt <- lm(y ~ z, subset = L < y & y < R)
abline(mt, lty = 2, col = "black")

## black observed points
s <- L < y & y < R
points(z[s], y[s])

## let's do as with OLS and normal x
set.seed(2)
N <- 100
x <- rnorm(N, 100, 5)
b <- .5
eps <- rnorm(N, 0, 5)
y <- 30 + b * x + eps
png(here("figures/fig2-regression_example.png"), 
    width = 7, height = 4, units="in", res=300)
par(mfrow = c(1, 2))
plot(x, y,
  col = "black",
  xlab = "covariate",
  ylab = "death age"
)
m <- lm(y ~ x)
abline(m, lty = 1, col = "black", lwd = 2)
title("True slope = 0.45")

## create plot
plot(x, y,
  col = "black",
  xlab = "covariate",
  ylab = "death age"
)

m <- lm(y ~ x)
abline(m, lty = 1, col = "black")
L <- 75
R <- 85
s <- L < y & y < R
mt <- lm(y ~ x, subset = s)
abline(mt, lty = 1, col = "red", lwd = 2)
abline(h = c(L, R), lty = 3, col = "red")
points(x[s], y[s], pch = 19, cex = .5, col = "red")
title("Truncated slope = 0.15", col.main = "red")
dev.off()
```
