---
title: "Figure 4 - CenSoc DMF 1910 Diagnostics"
author: Casey Breen
---

Summary: This .Rmd file contains code to replicate Figure 4 in "Mortality Modeling of Partially Observed Cohorts Using Administrative Death Records." 

Description of Figure 4: For CenSoc-DMF cohort of 1910: panel (a) shows a histogram of age of death for those who did and did not complete high school; the black curves shows our Gompertz-model based estimates of the number of deaths.  Panel (b) plots the model-based hazard ratio estimates (solid lines) vs. the inferred observed hazard ratios (dashed lines) for those who completed high school (blue) and those who did not (red).


```{r}
## library packages
library(gompertztrunc)
library(tidyverse)
library(data.table)
library(broom)
library(here)
source("helpers.R")
```


## We simulate without any cohort distinctions

```{r}
## read in censoc_dmf_v2_linked.csv
dmf <- fread(here("data/censoc_dmf_v2.1_linked_extract.csv"))

## filter to the year 1910
dmf_1910 <- dmf %>%
  filter(byear == 1910) %>%
  filter(link_abe_exact_conservative == 1) %>%
  filter(!EDUCD %in% c(0, 999)) %>%
  recode_education() %>%
  mutate(hs = case_when(
    educ_yrs >= 12 ~ "High School",
    TRUE ~ "No High School"
  )) %>%
  mutate(MARST = as.factor(MARST))
```

```{r}
## estimate gradient
gradient <- gompertz_mle(formula = death_age ~ hs, data = dmf_1910, left_trunc = 1975, right_trunc = 2005)

## diagnostic plots of curve fitting
model_plot <- diagnostic_plot(
  object = gradient, data = dmf_1910,
  covar = "hs", death_var = "death_age", xlim = c(50, 100)
) +
  ggplot2::scale_color_manual(
    name = "",
    values = c(Modeled = "black")
  ) +
  labs(title = "CenSoc-DMF Cohort of 1910")

## diagnostic plots of model hazards
model_plot_hazards <- diagnostic_plot_hazard(
  object = gradient, data = dmf_1910,
  covar = "hs", death_var = "death_age", xlim = c(65, 95)
)

## combine plots
combined_hazards <- cowplot::plot_grid(model_plot, model_plot_hazards, labels = "auto", nrow = 2)

## save ggplots
ggsave(
  plot = combined_hazards,
  filename = here("figures/fig4-histograms_and_hazards.png"),
  width = 10, height = 11
)
```
