---
title: "R Notebook"
author: Casey Breen 
---

Summary: This .Rmd file contains code to replicate Figure 5 in "Mortality Modeling of Partially Observed Cohorts Using Administrative Death Records." 

Description of Figure 5: For CenSoc-DMF cohort of 1910: panel (a) shows a histogram of age of death for those who did and did not complete high school; the black curves shows our Gompertz-model based estimates of the number of deaths.  Panel (b) plots the model-based hazard ratio estimates (solid lines) vs. the inferred observed hazard ratios (dashed lines) for those who completed high school (blue) and those who did not (red).

```{r}
## library packages
library(data.table)
library(tidyverse)
library(broom)
library(socviz)
library(cowplot)
library(ggsci)
library(here)
source("helpers.R")
```



```{r}
## read in prelinked CenSoc datasets and filter to "conservative" matches
dmf <- fread(here("data/censoc_dmf_v2.1_linked_extract.csv")) %>%
  filter(link_abe_exact_conservative == 1)

numident <- fread(here("data/censoc_numident_v2.1_linked_extract.csv")) %>%
  filter(link_abe_exact_conservative == 1)
```



```{r}
## recode dmf education variable
dmf <- dmf %>%
  recode_education() %>% ## function from "helper functions"
  filter(!is.na(educ_yrs))

## recode dmf education variable
dmf_1988_2005 <- dmf %>%
  filter(dyear %in% c(1988:2005))

## recode numident education variable
numident <- numident %>%
  recode_education() %>%
  filter(!is.na(educ_yrs))
```


```{r}
## linear model on full dmf
dmf_1910_lm <- tidy(lm(death_age ~ educ_yrs + as.factor(byear), weights = weight_conservative, data = dmf %>% filter(byear %in% 1905:1914) %>% sample_n(100000))) %>% mutate(dataset = "DMF (1975-2005)")

## linear model of dmf restricted to 1988-2005
dmf_1910_restrict_lm <- tidy(lm(death_age ~ educ_yrs + as.factor(byear), weights = weight_conservative, data = dmf_1988_2005 %>% filter(byear %in% 1905:1914) %>% sample_n(100000))) %>% mutate(dataset = "DMF (1988-2005)")

## linear model on numident
numident_1910_lm <- tidy(lm(death_age ~ educ_yrs + as.factor(byear), weights = weight_conservative, data = numident %>% filter(byear %in% c(1905:1914) & sex == 1) %>% sample_n(100000))) %>% mutate(dataset = "Numident (1988-2005)")
```


```{r}
## Gompertztrunc for DMF
dmf_1910 <- gompertztrunc::gompertz_mle(death_age ~ educ_yrs, data = dmf %>% filter(byear %in% 1905:1914), weights = weight_conservative)

dmf_1910_e65 <- gompertztrunc::convert_hazards_to_ex(dmf_1910$results) %>%
  mutate(dataset = "DMF (1975-2005)")

## Gompertztrunc for dmf truncated (1988-2005)
dmf_1910_1988_2005 <- gompertztrunc::gompertz_mle(death_age ~ educ_yrs, data = dmf_1988_2005 %>% filter(byear %in% 1905:1914), left_trunc = 1988, right_trunc = 2005, weights = weight_conservative)

dmf_1910_1988_2005_e65 <- gompertztrunc::convert_hazards_to_ex(dmf_1910_1988_2005$results) %>%
  mutate(dataset = "DMF (1988-2005)")

## Gompertztrunc for numident 1988-2005
numident_1910 <- gompertztrunc::gompertz_mle(death_age ~ educ_yrs, data = numident %>% filter(byear %in% 1905:1914 & sex == 1), left_trunc = 1988, right_trunc = 2005, weights = weight_conservative)

numident_1910_e65 <- gompertztrunc::convert_hazards_to_ex(numident_1910$results) %>%
  mutate(dataset = "Numident (1988-2005)")
```


```{r}
## combine linear model results
lm_results <- dmf_1910_lm %>%
  bind_rows(dmf_1910_restrict_lm) %>%
  bind_rows(numident_1910_lm) %>%
  filter(term == "educ_yrs") %>%
  mutate(method = "Regression on Age of Death") %>%
  mutate(
    e65 = estimate,
    e65_lower = estimate - 1.96 * std.error,
    e65_upper = estimate + 1.96 * std.error
  )

## combine gompertz results
gompertz_results <- dmf_1910_e65 %>%
  bind_rows(dmf_1910_1988_2005_e65) %>%
  bind_rows(numident_1910_e65) %>%
  mutate(method = "Parametric Gompertz (Accounting for Truncation)")

## create education plot
education_assocation_plot <- gompertz_results %>%
  bind_rows(lm_results) %>%
  ggplot(aes(x = dataset, y = e65, ymin = e65_lower, ymax = e65_upper, color = method, shape = method)) +
  geom_pointrange(position = position_dodge2(width = .1)) +
  theme_cowplot() +
  scale_shape_manual(values = c(1, 2)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  labs(
    x = "",
    y = "estimate",
    title = "Association Between Education (Years) and Longevity",
    subtitle = "Men, birth cohorts of 1910"
  ) +
  ggsci::scale_color_lancet() +
  ylim(0, .4)

## save plot
ggsave(
  plot = education_assocation_plot,
  filename = here("figures/fig5-education_association.png"),
  height = 6, width = 8
)
```
